Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-11 article/article.tex
Commit 2024-06-11 20:14

Figure\nobreakspace{}\ref{fig:P-energy} shows the time series of the
yearly number of CE cases (each with duration of one minute), the yearly
mean OIR and the yearly excess irradiation for the period 1993 -- 2023,
together with corresponding linear trends. All three quantities show
increasing trends, most pronounced for the frequency of occurrence
(\(45.6\pm 21.9\,\text{cases}/\text{year}\)) and the excess irradiation
(\(116.9\pm 67.8\,\text{kJ}/\text{year}\)), which are also statistically
significant. In contrast the trend of the yearly mean OIR is negligible
(\(0.1\pm 0.1\,\,\text{W}/\text{m}^2/\text{year}\)) and of no
statistical significance.

Although, the previous result is closer to the raw data, we preferred to
present the annual statistics, that give a more clear picture about the
long term CE trends. Hence, we calculated the annual values from the
one-minute measurements. Despite the very weak statistical significance
of the mean OIR trend, the annual number of CE occurrences shows a
steady increase of \(45.6\pm 21.9\,\text{cases}/\text{year}\)
(Figure\nobreakspace{}\ref{fig:P-energy-N}). Subsequently, the mean
annual energy of the CE events is increasing with a rate of
\(116.9\pm 67.8\,\text{kJ}/\text{year}\)
(Figure~\ref{fig:P-energy-sum}), witch follows the trend of the number
of CE occurrences. We have to note that the energy related to the CE
events can not be directly linked with the total energy balance of the
atmosphere. The net solar radiation of the region is not increased, but
rather redistributed through the CE events. This also is depicted by the
``instantaneous'' values of irradiance (average over one minute), that
can exceed the equivalent clear sky irradiance to a considerable level.

@@

\hypertarget{cloud-enhancements-climatology}{%
\subsection{Cloud enhancements
climatology}\label{cloud-enhancements-climatology}}

Another interesting aspect of the CE cases occurrences, is their
seasonal cycle. In
Figure\nobreakspace{}\ref{fig:relative-month-occurrences}, we have the
monthly box plot (whisker plot), where the values have been normalized
by the highest median value, that occurs on June. Although the number of
occurrences has a wide spread throughout the study period, the most
active period of CE occurrences is during May and June. During the
Winter (December -- February) the CE cases are about \(25\,\%\) of the
maximum. For the rest months, the occurrences gradually ramp between the
maximum and minimum. This seasonality is a combinations of the clouds
occurrences, and the relevant position of the sun during each season of
the year. Unfortunately, the lack of detailed data on cloud formation,
type and location is not allowing further analysis.

>>>>

Figure\nobreakspace{}\ref{fig:P-energy} shows the time series of the
yearly number of CE cases (each with duration of one minute), the yearly
mean OIR and the yearly excess irradiation for the period 1993 -- 2023,
together with corresponding linear trends. All three quantities show
increasing trends, most pronounced for the frequency of occurrence
(\(45.6\pm 21.9\,\text{cases}/\text{year}\)) and the excess irradiation
(\(116.9\pm 67.8\,\text{kJ}/\text{year}\)), which are also statistically
significant. In contrast the trend of the yearly mean OIR is negligible
(\(0.1\pm 0.1\,\text{W}/\text{m}^2/\text{year}\)) and of no statistical
significance. The average OIR for the entire period is about
\(39.9\,\text{W}/\text{m}^2\) with standard deviation of
\(2.7\,\text{W}/\text{m}^2\). The interannual variability of the data
about the trend lines is quite large. Furthermore, it tends to increase
with time (at least for the quantities of panels b and c), suggesting a
significant variability in cloud patterns over the area, possibly
associated to changes in climate.

We have to note that the excess irradiation related to the CE events can
not be directly linked to the total energy balance of the atmosphere.
The net solar radiation of the region is not increased, but is rather
redistributed through the CE events. This is also depicted by the ECE
irradiance values, which can exceed the equivalent clear sky irradiance
by a significant amount.

@@

\hypertarget{climatology-of-cloud-enhancement-events}{%
\subsection{Climatology of cloud enhancement
events}\label{climatology-of-cloud-enhancement-events}}

Another interesting aspect of the CE events, is their distribution
within the year. In
Figure\nobreakspace{}\ref{fig:relative-month-occurrences}, we have the
monthly box plot (whisker plot), where the values have been normalized
by the highest median value, that occurs on June. Although the number of
occurrences has a wide spread throughout the study period, the most
active period of CE occurrences is during May and June. During the
Winter (December -- February) the CE cases are about \(25\,\%\) of the
maximum. For the rest months, the occurrences gradually ramp between the
maximum and minimum. This seasonality is a combinations of the clouds
occurrences, and the relevant position of the sun during each season of
the year. Unfortunately, the lack of detailed data on cloud formation,
type and location is not allowing further analysis.
